Approximation as Differentiation

### 2025-06-18 main/first-order.tex
\downset.

For a first-order language with a primitive type $\syn{nat}$ and primitive operations such as $+$, the
``leaves'' of the approximation lattices would provided by a model of $\syn{nat}$ and its operations in
$\Fam(\LatGal)$. For example, we might assign $\syn{nat}$ to the object $(\Nat,n \mapsto \{n, \bot\})$, where
the family component is the two-point lattice indicating whether a given natural number is fully present ($n$)
or fully absent ($\bot$) from the approximation, as shown in the example in \secref{introduction}, and we
might then assign $+$ to the morphism $(\hat+,\partial(\hat+))$, with $\partial(\hat+)$ defined as
\begin{displaymath}
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \partial(\hat+)&n&m&=&n \mathbin{\hat+} m \\
      \partial(\hat+)&\bot&\_&=&\bot \\
      \partial(\hat+)&\_&\bot&=&\bot
    \end{array}
\end{displaymath}
These elementary approximations would then be combined by the interpretation of composite types like sums and
products, providing suitable derived approximation structures for composite values.

>>>>

For a first-order language with a primitive type $\syn{nat}$ and primitive operations such as $+$, the
language would provide a model of $\syn{nat}$ and its operations in $\Fam(\LatGal)$. For example, we might
interpret $\syn{nat}$ to the object $(\Nat,n \mapsto \downset{n})$, where the family component assigns to
every natural number $n$ the two-point lattice $\downset{n} = \{n, \bot\}$ indicating whether $n$ is fully
present ($n$) or fully absent ($\bot$) from an approximation of $n$, as shown in the example in
\secref{introduction}. We might then interpret $+$ as the morphism $(+,\partial+)$, with $+$ here meaning the
usual addition operator, and $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as
\begin{displaymath}
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \partial+_{n,m}&n&m&=&n + m \\
      \partial+_{n,m}&\bot&\_&=&\bot \\
      \partial+_{n,m}&\_&\bot&=&\bot
    \end{array}
\end{displaymath}
These elementary approximations would then be combined by the interpretation of composite types like sums and
products, providing suitable derived approximation structures for composite values.
